Self Referencing Sequences

### 2014-11-04 draft.tex
Edits to the first 3 pages

The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). Thus, this sequence is \emph{self-referential}.

Observe that in this case it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. However ambiguity exists in a sequence generated by $(1,2,3)$.
If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
\begin{equation}
1, 3, 3, 3, 2, 2, 2, \dots
\end{equation}
\begin{equation}
1, 2, 2, 1, 1, 3, 1, \dots\\
\end{equation}
are in some sense ``self-referencing'' sequences.
Each time we need to specify a new block, we have a choice of two elements. We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence. 
For example, $(1, 2, 3 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.

With this rule, the self-referencing sequence generated by $(1,2,3)$ that begins with a $1$ is
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
\end{figure}

@@

Our main results are:
\begin{itemize}
\item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
\item Some characterizations of sequences with the same generating cycle.
\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequencem if a limiting density exists.
\item The limiting densities of the self-referencing sequences of  $(1,3)$ if such densities exists.
\end{itemize}

In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. We consider how the starting number for the sequence affects the long term behavior of the sequence. In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

In Section~\ref{iterativeprocess}, we take a different approach to analyzing these sequences by showing that there is an alternate way of generating them with an iterative process. In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules. We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets. Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.

@@

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number.

@@

In this section, we consider a different definition of equivalence. Consider, for example, sequences generated by $C = (1,2,3)$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
2,2,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases.

Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}. To formalize, this, we say that two sequences $(a_i)$ and $(b_i)$ are equivalent if there exists some integers $n,k \geq 0$ such that
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}
for all positive integers $i$. This definition is saying that the sequences are eventually equivalent.

\begin{conjecture} Any sequence generated by a set $S$ is equivalent to any other sequence generated by the same set $S$.
\end{conjecture}

This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (1,2,3)$. For generating cycles with larger numbers, we have been unable to verify or refute the conjecture.. However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.

>>>>

The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). We call such a sequence \emph{self-referential}.

Observe that it was unambiguous which number a new block begins with. After a block of 1's, the next block must be a block of 2's and vice versa. However if our sequence was composed of $1$'s, $2'$s and $3'$s there would be ambiguity over which number a new block should begin with. 
For example if the first number is 1, then we could make the second number either a 2, or a 3. In particular, both:
\begin{equation}
1, 3, 3, 3, 2, 2, 2, \dots
\end{equation}
\begin{equation}
1, 2, 2, 1, 1, 3, 1, \dots\\
\end{equation}
are in some sense ``self-referencing'' sequences.
Each time we need to specify a new block, we have a choice of two elements. We would like to avoid having a large number of choices in the generation of the sequence, and so we will use the convention that the order in which the elements are presented gives the order they are used in the sequence. 
For example, $(123 )$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$.

With this rule, the self-referencing sequence generated by $(123)$ that begins with a $1$ is
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(123)$ broken up into contiguous blocks.}
\end{figure}

@@

Our main results are:
\begin{itemize}
\item The self-referencing sequence for a given generating cycle is uniquely determined by its first number.
\item Some characterizations of sequences with the same generating cycle.
\item For some generating cycles there exists an equivalent formulation of self-referencing sequences using substitution rules
\item If a generating cycle has substitution rules then one can compute the limiting density for the numbers in the sequence if a limiting density exists.
\item The limiting densities of the self-referencing sequences of  $(13)$ if such densities exists.
\end{itemize}

In Section~\ref{determinism} we show that if we are given a generating cycle and a starting number for the sequence, the sequence is uniquely determined. In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating cycle, and not the starting number.

In Section~\ref{iterativeprocess}, we show that there is an alternate way of generating them with an iterative process. In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules. We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating cycles. Unfortunately, we find that the open problem of determining the density of the sequences generated by $(12)$ is hard to solve using our methods.

@@

We claim that a self-referencing sequence is uniquely determined by the generating cycle and the starting number.

@@

In this section, we consider long term equivalence of sequences. We introduce a definition for this
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}
for all positive integers $i$. 
\end{definition}

 Consider, for example, sequences generated by $C = (123)$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
2,2,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases.

\begin{conjecture} Any sequence generated by a cycle $C$ is equivalent to any other sequence generated by the same cycle $C$.
\end{conjecture}

This conjecture is partly motivated by the fact that the property holds for self-referencing sequences from the generating cycle $C= (123)$. For generating cycles with larger numbers, we have been unable to verify or refute the conjecture.. However, in Section~\ref{density}, we show that for some generating cycles the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.
